---
title: It's Parking Time!
author: geekyboy88
date: '2017-12-24'
slug: its-parking-time
output:
  html_document:
    toc: true
categories: []
tags: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, comment=NA, message=FALSE)
```

### Open Data in Singapore

Hi again! For my second post, I would like to focus on exploring the open data sources available at [data.gov.sg](https://www.data.gov.sg/). But should one click on the link that I just provided and take a look at the available data sources, one would find an immense number of data sets available, and wouldn't necessarily know where to begin!

### Holiday Season!

As we approach the festive season, with Christmas, New Year, and Chinese New Year coming in the following months, 

### Load Packages

First 

---
```{r loadpackages, results='asis'}
library(magrittr)
library(jsonlite)
library(dplyr)
library(ggmap)
```


```{r}
library(MASS)
data(cats)
plot(Hwt ~ Bwt, data=cats, xlab="Body weight (kg)", ylab="Heart weight (g)")
```

### Reading in the Data

Next, let's read in the data into our working environment using the provided __API Key__.

```{r readdata, echo = T}
data = fromJSON("https://data.gov.sg/api/action/datastore_search?resource_id=139a3035-e624-4f56-b63f-89ae28d4ae4c&limit=4000")
data = data$result$records



```










_Note: This post contains information on [HDB Carpark Information](https://data.gov.sg/dataset/hdb-carpark-information?view_id=398e65ae-e2cb-4312-8651-6e65d6f19ed1&resource_id=139a3035-e624-4f56-b63f-89ae28d4ae4c) accessed [today's date](https://www.google.com.sg/search?q=today%27s+date&oq=today%27s+date&aqs=chrome..69i57.1121j0j7&sourceid=chrome&ie=UTF-8) which is made available under the terms of the [Singapore Open Data Licence version 1.0](https://data.gov.sg/open-data-licence)._
