---
title: It's Parking Time!
author: geekyboy88
date: '2018-01-06'
categories: []
tags: []
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

### Open Data in Singapore

Hi again! For my second post, I would like to focus on exploring the open data sources available at [data.gov.sg](https://www.data.gov.sg/). But should one click on the link that I just provided and take a look at the available data sources, one would find an immense number of data sets available, and wouldn't necessarily know where to begin!

### Holiday Season!

I thought, as we approach the festive season, with Christmas, New Year, and Chinese New Year coming in the following months, a good place to start would be the HDB Carparks dataset. Given that close to 80% of Singapore residents live in HDB flat, it is quite likely that you will be visiting one of your many relatives there. For those of you who are driving, you may be interested to find out where you can get free parking, etc.

### Load Packages

First, let's load some package that we will need.

```{r loadpackages, message=FALSE, warning=FALSE}
library(magrittr)
library(jsonlite)
library(knitr)
library(dplyr)
library(ggmap)
```

### Reading in the Data

Next, let's read in the data into our working environment using the provided __API Key__. After we read in the data (_in JSON format_), we realise that the API provides data, as well as the metadata of the dataset. However, we will only keep the actual dataset because that is what we are interested in. 

```{r readdata, echo = T, message=FALSE, warning=FALSE}
data = fromJSON("https://data.gov.sg/api/action/datastore_search?resource_id=139a3035-e624-4f56-b63f-89ae28d4ae4c&limit=2502")
data = data$result$records
```

### Exploring the dataset

Let's explore this dataset further. Taking a look below, we can see a total of 10 data fields, which gives us information of the address and location of the carpark, what type of carpark it is, the type of parking system, and whether you can park there for free or during the night. In total, there are 2052 HDB carparks as at the date of this post.

```{r exploredata, echo=F, message=NA}
knitr::kable(head(data))
nrow(data)
```

## Where are the locations of the carparks?



```{r location}

```



#### CURRENTLY UNDER CONSTRUCTION

_Stay Tune_!




_Note: This post contains information on [HDB Carpark Information](https://data.gov.sg/dataset/hdb-carpark-information?view_id=398e65ae-e2cb-4312-8651-6e65d6f19ed1&resource_id=139a3035-e624-4f56-b63f-89ae28d4ae4c) accessed [today's date](https://www.google.com.sg/search?q=today%27s+date&oq=today%27s+date&aqs=chrome..69i57.1121j0j7&sourceid=chrome&ie=UTF-8) which is made available under the terms of the [Singapore Open Data Licence version 1.0](https://data.gov.sg/open-data-licence)._
