---
title: Soccer and Discipline
author: geekyboy88
date: '2018-01-07'
slug: soccer-and-discipline
categories:
  - R
  - soccer
  - Analytics
tags:
  - R
  - R Markdown
  - soccer
  - analytics
---

<style type="text/css">

body, td {
   font-size: 14px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
</style>

### Soccer and Discipline

Hi everyone again. For my third post, I have decided to look at some soccer data that was recently uploaded at Kaggle. This is a mega dataset comprising close to 900k records of events across a number of major European leagues over the course of about 5 seasons. The __great__ thing about events data is that it gives you a really intricate flavour of what happens in the match. 

#### Packages

Again, let's start off with the basic steps of loading the required packages.

```{r loadpackages, message=FALSE, warning=FALSE}
library(magrittr)
library(RColorBrewer)
library(ggplot2)
library(knitr)
library(scales)
library(dplyr)
```

#### Data

After loading in the dataset, let's have a quick look at the information that each dataset contains.

First, we'll take a look at the events information. 

```{r loaddata, message=FALSE, warning=FALSE}
events = read.csv("../../static/data/events1.csv", stringsAsFactors = F)
ginf = read.csv("../../static/data/ginf.csv", stringsAsFactors = F)

knitr::kable(head(events))
```

And next, we'll look at the data available in the matches dataset.

```{r viewginf, message=FALSE, warning=FALSE}
knitr::kable(head(ginf))
```

Since its my first time working on this dataset, I'll focus only on the English Premier League (EPL), just because it is the most watched league in Singapore. What I did was to first merge in the match information to the events data, and subsequently filter only for events that happened during EPL matches. Also, since we aren't doing comparisons across leagues, I have also filtered down to the 2015/2016 season (_yes, that season when Leicester City shocked? the world by becoming EPL champions_).

```{r mergedata, message=FALSE, warning=FALSE}
events %<>% merge(ginf[,c(1, 4:9)], by = "id_odsp", all.x = T)
rm(ginf)
```

### Disciplinary Records

For my first post, I wanted to focus on the other side of soccer that is typically not talked about as often. Typically, any commentary or analysis about soccer looks at __goals__ and/or __goal chances__, and the disciplinary side of things tend to be overlooked. However, I believe that there are interesting insights that can be taken away, as _red cards_, _yellow cards_, or _fouls_ have the potential to be __game-changing__ moments in the match. 

For a start, let's attempt to find out which of these EPL teams tend to have the best / worst discplinary records.

```{r cards, message=FALSE, warning=FALSE}
events %<>%
  filter(event_type == 4 | event_type == 6 | event_type == 3) %>%
  mutate(event_type = ifelse(event_type == 4, "Yellow Card",
                      ifelse(event_type == 6, "Red Card", "Foul")))
```

### Fouls

We start off by looking at the teams with the most fouls. _Well well well_, no surprise to see here that __Manchester United__ is top of the fouls league (_given the over-emphasis of breaking up play rather than making play in that season under Louis Van Gaal_). Following closely behind is __Crystal Palace__, __Tottenham Hotspur__ and __Watford__. At the other end of the table, we see that __Everton__ is the "_cleanest_" team with less than 300 fouls across the entire season.

```{r fouls, warnings=FALSE, message=FALSE}
events %>%
  filter(event_type == "Foul") %>%
  group_by(event_team) %>%
  summarise(num = n()) %>%
  ggplot(aes(x = reorder(event_team, num), y = num, fill = num)) +
  geom_bar(position = position_dodge(), stat ="identity") +
  scale_fill_gradient(low = "magenta1", high = "maroon") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "EPL team", y = "Number of Fouls", title = "Number of Fouls by EPL Teams") +
  coord_flip()
```

Similarly, we can do this for yellow cards and red cards. Taking in all the information across the 3 charts, we can rougly make some rough deductions. 

1. __Petty fouls__ - Having the most number of fouls doesn't necessarily mean more yellow cards; most of these fouls often result in a free-kick to the opposition and nothing more. 
2. __A moment of red?__ - We also don't observe any strong correlation between the number of fouls, yellow cards and red cards. For example, we see that while __West Ham__ is clearly in the lead with 5 red cards, they are clearly near the bottom of the table for number of fouls and yellow cards. This could suggest that straight red cards are likely to be borne out of a moment of recklessness, anger, or folly. 

```{r yellowcard}
events %>%
  filter(event_type == "Yellow Card") %>%
  group_by(event_team) %>%
  summarise(num = n()) %>%
  ggplot(aes(x = reorder(event_team, num), y = num, fill = num)) +
  geom_bar(position = position_dodge(), stat ="identity") +
  scale_fill_gradient(low = "yellow4", high = "yellow1") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "EPL team", y = "Number of Yellow Cards", title = "Number of Yellow Cards by EPL Teams") +
  coord_flip()
```


```{r redcard}
events %>%
  filter(event_type == "Red Card") %>%
  group_by(event_team) %>%
  summarise(num = n()) %>%
  ggplot(aes(x = reorder(event_team, num), y = num, fill = num)) +
  geom_bar(position = position_dodge(), stat ="identity") +
  scale_fill_gradient(low = "red4", high = "red1") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "EPL team", y = "Number of Red Cards", title = "Number of Red Cards by EPL Teams") +
  coord_flip()
```

So... perhaps approaching disciplinary records from a team perspective may not be ideal. Let's try it from the player's perspective instead. For this analysis, I decided to create a discplinary score for each player. A yellow card gives you 1 point, while a red card gives you 3 points. Of course this is all just arbitary, and can be revised if your opinion differs. At the top of the discplinary record list is __Gary O'Neil__. Not exactly who I would expect to be topping the charts for worst disciplinary records, but yes, __Gary O'Neil__ is our man for that season.

```{r player}
events %>%
  mutate(discplinary_score = ifelse(event_type == "Yellow Card", 1,
                             ifelse(event_type == "Red Card", 3, 0))) %>%
  group_by(player) %>%
  summarise(Total_Score = sum(discplinary_score),
            Num_Yellow = sum(ifelse(event_type == "Yellow Card", 1, 0)),
            Num_Red = sum(ifelse(event_type == "Red Card", 1, 0))) %>%
  arrange(-Total_Score) %>%
  head(20)


```

### Petty Fouls

In the previous section, we looked at discipline from the angle of receiving some form of __punishment__, e.g. through yellow or red cards. However, in normal games, players are often given a certain degree of leeway (_unless its a horrific foul_) before they get their first booking (_often a yellow card_). In this section, I decided to bring the fouls committed back into the fold, and try and identify players who have a typically good knack of committing a number of fouls without incurring any form of punishment.

To do so, I first aggregated, for each player playing in a match, what is the number of fouls committed and yellow cards received in that match. Using these two information, I constructed a fouls to yellow cards ratio _(e.g. if Nemanja Matic commits 4 fouls and receives 1 yellow card his ratio is __4:1__)_ I subsequently counted the number of times each player has this particular fouls to yellow cards ratio.

```{r foulspergame}
fouls_per_game = events %>% 
  group_by(id_odsp, player, event_team) %>%
  summarise(num_foul = sum(ifelse(event_type == "Foul", 1, 0)) +
              sum(ifelse(event_type == "Yellow Card", 1, 0)),
            num_yellow = sum(ifelse(event_type == "Yellow Card", 1, 0))) %>%
  mutate(fouls_to_yellow = paste0(num_foul, ":", num_yellow))

player_ratio = fouls_per_game %>%
  group_by(player, fouls_to_yellow) %>%
  summarise(count = n())
```

With this, let's take a look at the top 10 fouls to yellow cards ratio and the players associated with it. From the table below, we can tell that Nemanja Matic is the true champion in committing the most number of fouls and getting away with it, as he had on one occasion, committed 8 fouls in a match without incurring a yellow card. Other players who have committed a number of fouls in the similar range have, _unfortunately_, found themselves punished with a yellow card.

_Note: The number of fouls include the foul that led to a yellow card._

```{r topfoultoyellow}
kable(head(player_ratio %>% 
             arrange(desc(fouls_to_yellow)), 15))
```

### Likelihood of seeing yellow

So this then leads to the question: what is the likelihood of you getting a yellow card? Well, to comprehensively answer this question, I thought we could start off by imagining ourselves as a player starting a match with a clean slate. At some point in time, you go in for a tackle (_not knowing if it will be a clean, clumsy or reckless tackle_), and right at that point in time when the referee blows his whistle (_meaning to say that the tackle you made was definitely not a clean one_) you start to wonder what is the chance that you will be given a yellow card. 

To find out what is the __probability of you getting a yellow card on your first foul__, we first include every player who has ever committed at least 1 foul into our sample.

_Note: For simplicity, assume that the yellow card is given to the player on his last foul, i.e for a foul to yellow ratio of 4:1, the yellow card is given to the player on his 4th foul._

Based on the chart below, overall, condition on committing a foul (_or getting called for committing a foul by the referee, given that you could technically get away with committing one_), you have a __6.7%__ chance of getting a yellow card for your first foul. This is quite consistent with the perception that players usually get away with their first fouls.

We could also inteprete this another way, by taking into account the heterogeneous nature of fouls. Fouls can either be soft, clumsy, slightly late, very late, or reckless, etc. If we then assume then that referees have eagle eye and consistently award yellow cards for fouls that are deserving of one, this __6.7%__ is actually, for players who committed a foul, the probability of them committing a reckless one.

```{r likelihood1}
player_ratio %>%
  filter(fouls_to_yellow != "0:0" & substr(fouls_to_yellow,3,3) != "2") %>%
  mutate(group = ifelse(fouls_to_yellow == "1:1", "Yellow on First Foul", "No Yellow on First Foul")) %>%
  group_by(group) %>%
  summarise(count = n()) %>%
    mutate(cumulative = cumsum(count),
         midpoint =  sum(count) - (cumulative - count/2)) %>%
  ggplot(aes(x = "", y = count, fill = group)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values = c("grey95","yellow")) +
  coord_polar("y", start = 0) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "", y = "", fill = "Legend", title = "Probability of getting a yellow card \non your first foul") +
  geom_text(aes(y = midpoint, label = percent(count/sum(count))), size = 5)
```

If one foul is all you are going to commit during the entire match, then the above analysis should be sufficient for you. However, if you had committed the first foul early in the match, then you would probably want to read on to find out if you again hear the referee's whistle after going in for a tackle. 

To estimate the probability of getting a yellow card, we now sub-sample to exclude those who committed less than 2 fouls in a match. 

Based on the chart below, we find that the probability of you getting a yellow card on your second foul is significantly higher at __22%__. 

```{r likelihood2}
player_ratio %>%
  filter(as.numeric(substr(fouls_to_yellow,1,1)) > 1 & substr(fouls_to_yellow,3,3) != "2") %>%
  mutate(group = ifelse(fouls_to_yellow == "2:1", "Yellow on Second Foul", "No Yellow on Second Foul")) %>%
  group_by(group) %>%
  summarise(count = n()) %>%
    mutate(cumulative = cumsum(count),
         midpoint =  sum(count) - (cumulative - count/2)) %>%
  ggplot(aes(x = "", y = count, fill = group)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values = c("grey85","yellow")) +
  coord_polar("y", start = 0) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "", y = "", fill = "Legend", title = "Probability of getting a yellow card \non your second foul") +
  geom_text(aes(y = midpoint, label = percent(count/sum(count))), size = 5)
```

We repeat the same for the third (__30.3%__) and fourth (__39.2%__) fouls. Not surprisingly, the probabilities are increasing, which is not surprising because your name becomes more salient with the referee with each foul, and there comes a stage where the referee utimately decides to make you see yellow. However, do note that the probabilities cannot be compared with reference to each other, each situation is evaluated on its on (_conditional probabilites_).

```{r likelihood3}
player_ratio %>%
  filter(as.numeric(substr(fouls_to_yellow,1,1)) > 2 & substr(fouls_to_yellow,3,3) != "2") %>%
  mutate(group = ifelse(fouls_to_yellow == "3:1", "Yellow on Third Foul", "No Yellow on Third Foul")) %>%
  group_by(group) %>%
  summarise(count = n()) %>%
    mutate(cumulative = cumsum(count),
         midpoint =  sum(count) - (cumulative - count/2)) %>%
  ggplot(aes(x = "", y = count, fill = group)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values = c("grey75","yellow")) +
  coord_polar("y", start = 0) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "", y = "", fill = "Legend", title = "Probability of getting a yellow card \non your third foul") +
  geom_text(aes(y = midpoint, label = percent(count/sum(count))), size = 5)
```


```{r likelihood4}
player_ratio %>%
  filter(as.numeric(substr(fouls_to_yellow,1,1)) > 3 & substr(fouls_to_yellow,3,3) != "2") %>%
  mutate(group = ifelse(fouls_to_yellow == "4:1", "Yellow on Fourth Foul", "No Yellow on Fourth Foul")) %>%
  group_by(group) %>%
  summarise(count = n()) %>%
    mutate(cumulative = cumsum(count),
         midpoint =  sum(count) - (cumulative - count/2)) %>%
  ggplot(aes(x = "", y = count, fill = group)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values = c("grey65","yellow")) +
  coord_polar("y", start = 0) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "", y = "", fill = "Legend", title = "Probability of getting a yellow card \non your fourth foul") +
  geom_text(aes(y = midpoint, label = percent(count/sum(count))), size = 5)
```

Hope you enjoy this post and look forward to the next post! If you have any thoughts or comments, do hold on to that thought with you for now __or__ if you can share it with me if you can find a way to contact me. I'll try and create a comments section soon!



Data source obtained from [Kaggle](https://www.kaggle.com/). Credit to Alin Secareanu for compiling this amazing dataset. Data can be accessed from this [link](https://www.kaggle.com/secareanualin/football-events).