---
title: Soccer and Anger Management
author: geekyboy88
date: '2018-01-07'
slug: soccer-and-anger-management
categories:
  - R
  - soccer
  - Analytics
tags:
  - R
  - R Markdown
  - soccer
  - analytics
---

<style type="text/css">

body, td {
   font-size: 14px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
</style>

### Soccer and Anger Management

Hi everyone again. For my third post, I have decided to look at some soccer data that was recently uploaded at Kaggle. This is a mega dataset comprising close to 900k records of events across a number of major European leagues over the course of about 5 seasons. The __great__ thing about events data is that it gives you a really intricate flavour of what happens in the match. 

#### Packages

Again, let's start off with the basic steps of loading the required packages.

```{r loadpackages, message=FALSE, warning=FALSE}
library(magrittr)
library(RColorBrewer)
library(ggplot2)
library(knitr)
library(dplyr)
```

#### Data

After loading in the dataset, let's have a quick look at the information that each dataset contains.

First, we'll take a look at the events information. 

```{r loaddata, message=FALSE, warning=FALSE}
events = read.csv("../../static/data/events1.csv", stringsAsFactors = F)
ginf = read.csv("../../static/data/ginf.csv", stringsAsFactors = F)

knitr::kable(head(events))
```

And next, we'll look at the data available in the matches dataset.

```{r viewginf, message=FALSE, warning=FALSE}
knitr::kable(head(ginf))
```

Since its my first time working on this dataset, I'll focus only on the English Premier League (EPL), just because it is the most watched league in Singapore. What I did was to first merge in the match information to the events data, and subsequently filter only for events that happened during EPL matches. Also, since we aren't doing comparisons across leagues, I have also filtered down to the 2015/2016 season (_yes, that season when Leicester City shocked? the world by becoming EPL champions_).

```{r mergedata, message=FALSE, warning=FALSE}
events %<>% merge(ginf[,c(1, 4:9)], by = "id_odsp", all.x = T)
rm(ginf)
```

### Disciplinary Records

For my first post, I wanted to focus on the other side of soccer that is typically not talked about as often. Typically, any commentary or analysis about soccer looks at __goals__ and/or __goal chances__, and the disciplinary side of things tend to be overlooked. However, I believe that there are interesting insights that can be taken away, as _red cards_, _yellow cards_, or _fouls_ have the potential to be __game-changing__ moments in the match. 

For a start, let's attempt to find out which of these EPL teams tend to have the best / worst discplinary records.

```{r cards, message=FALSE, warning=FALSE}
events %<>%
  filter(event_type == 4 | event_type == 6 | event_type == 3) %>%
  mutate(event_type = ifelse(event_type == 4, "Yellow Card",
                      ifelse(event_type == 6, "Red Card", "Foul")))
```

### Fouls

We start off by looking at the teams with the most fouls. _Well well well_, no surprise to see here that __Manchester United__ is top of the fouls league (_given the over-emphasis of breaking up play rather than making play in that season under Louis Van Gaal_). Following closely behind is __Crystal Palace__, __Tottenham Hotspur__ and __Watford__. At the other end of the table, we see that __Everton__ is the "_cleanest_" team with less than 300 fouls across the entire season.

```{r fouls, warnings=FALSE, message=FALSE}
events %>%
  filter(event_type == "Foul") %>%
  group_by(event_team) %>%
  summarise(num = n()) %>%
  ggplot(aes(x = reorder(event_team, num), y = num, fill = num)) +
  geom_bar(position = position_dodge(), stat ="identity") +
  scale_fill_gradient(low = "magenta1", high = "maroon") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "EPL team", y = "Number of Fouls", title = "Number of Fouls by EPL Teams") +
  coord_flip()
```

Similarly, we can do this for yellow cards and red cards. Taking in all the information across the 3 charts, we can rougly make some rough deductions. 

1. __Petty fouls__ - Having the most number of fouls doesn't necessarily mean more yellow cards; most of these fouls often result in a free-kick to the opposition and nothing more. 
2. __A moment of red?__ - We also don't observe any strong correlation between the number of fouls, yellow cards and red cards. For example, we see that while __West Ham__ is clearly in the lead with 5 red cards, they are clearly near the bottom of the table for number of fouls and yellow cards. This could suggest that straight red cards are likely to be borne out of a moment of recklessness, anger, or folly. 

```{r yellowcard}
events %>%
  filter(event_type == "Yellow Card") %>%
  group_by(event_team) %>%
  summarise(num = n()) %>%
  ggplot(aes(x = reorder(event_team, num), y = num, fill = num)) +
  geom_bar(position = position_dodge(), stat ="identity") +
  scale_fill_gradient(low = "yellow4", high = "yellow1") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "EPL team", y = "Number of Yellow Cards", title = "Number of Yellow Cards by EPL Teams") +
  coord_flip()
```


```{r redcard}
events %>%
  filter(event_type == "Red Card") %>%
  group_by(event_team) %>%
  summarise(num = n()) %>%
  ggplot(aes(x = reorder(event_team, num), y = num, fill = num)) +
  geom_bar(position = position_dodge(), stat ="identity") +
  scale_fill_gradient(low = "red4", high = "red1") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "EPL team", y = "Number of Red Cards", title = "Number of Red Cards by EPL Teams") +
  coord_flip()
```

So... perhaps approaching disciplinary records from a team perspective may not be ideal. Let's try it from the player's perspective instead. For this analysis, I decided to create a discplinary score for each player. A yellow card gives you 1 point, while a red card gives you 3 points. Of course this is all just arbitary, and can be revised if your opinion differs. At the top of the discplinary record list is __Gary O'Neil__. 

```{r player}
events %>%
  mutate(discplinary_score = ifelse(event_type == "Yellow Card", 1,
                             ifelse(event_type == "Red Card", 3, 0))) %>%
  group_by(player) %>%
  summarise(Total_Score = sum(discplinary_score),
            Num_Yellow = sum(ifelse(event_type == "Yellow Card", 1, 0)),
            Num_Red = sum(ifelse(event_type == "Red Card", 1, 0))) %>%
  arrange(-Total_Score) %>%
  head(20)


```




















### CURRENTLY UNDER CONSTRUCTION

Sorry, this post is still under construction and will be completed soon! Thank you for your patience! :)



Data source obtained from [Kaggle](https://www.kaggle.com/). Credit to Alin Secareanu for compiling this amazing dataset. Data can be accessed from this [link](https://www.kaggle.com/secareanualin/football-events).